Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-09-05 papers/thesis-kirstie/thesis.tex
more cleanups

Another way to find values for $\Xi(T)$ is to use the second virial coefficent $B_{2}$ from the equation of state expressed in terms of the virial expansion. 
\begin{equation}\label{virialequation}
    \frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) 
\end{equation}
where 
\begin{equation}
	B_{2}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}(r)}-1\right)d\vec{r} 
\end{equation}
The second virial coefficient can be derived from Euler's equation for the internal energy of a system, 
\begin{equation}U=TS-PV+\mu{N}\end{equation}
using the relations from statistical thermophysics which views the thermodynamic quantities as ensemble averages
\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
which gives
\begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
Expressing the Helmholtz free energy in terms of the pair potential $V(r)$ results in the equation of state expressed in terms of the virial expansion as given by Equation~\ref{virialequation}.

@@

\begin{figure}[h!]
    \centering
     % FIXME\includegraphics[height=7cm]{MaxwellDTC-Fig1.pdf}
    \caption{Identifying densities at the solid-liquid transition using Maxwell's Double Tangent Construction}
    \label{fig:MaxwellDT}
  \end{figure}

@@

The pressure at the transistion point can be obtained from the point of intersection that occurs when the Gibbs free energy per atom is plotted against the pressure for both the liquid and the solid at fixed temperature, as shown in Figure \ref{fig:GibbsvsP}. These curves are constructed from the slope and cooresponding y-intercept data collected from lines tangent along the curves on the the Helmholtz free energy per atom verses inverse density plots for both the liquid and the solid. 
\begin{figure}[h!]
    \centering
     % FIXME\includegraphics[height=7cm]{MaxwellDTC-Fig2.pdf}
    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
    \label{fig:GibbsvsP}
  \end{figure}
%CURVE FITTING! - put in methods/analysis

@@

\begin{figure}[h!]
    \centering
     % FIXME\includegraphics[height=7cm]{MaxwellDTC-Fig3.pdf}
    \caption{The liquid and solid densities at the point of transistion are found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
    \label{fig:Pvsinvn}
  \end{figure}

@@

A description of cDFT ordinarily begins with considering the Grand Free Energy defined as
\begin{equation}\Phi=F-\mu{N}\end{equation}
%\begin{equation}\int\frac{N}{V}\frac{V_{ext}}{N}dV\end{equation}
which can then be expressed in terms of a varrying number density profile $n(\vec r)$
%\begin{equation}\Phi=F-\int n(\vec{r})\phi{(\vec r)}d\vec{r} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu{N}\end{equation}
\begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
where $\phi=\frac{V_{ext}}{N}$, $V_{ext}(\vec r)$ is an externally applied potential energy, and
\begin{equation}F_{int}= F -\int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}  
%If $\mu=0$ and $V_{ext}=0$, then the Grand free energy $\Phi$ reduces to the Helmholz free energy $F$. %The second and third terms in Eq~\ref{GrandFE} are both functionals of $n(\vec{r})$, so it remains to show that $F$, and thus $\Phi$, is a functional of $n(\vec{r})$.
The Grand free energy is applicable to systems with a varying number of particles and fixed chemical potential, temperature and volume. But this paper will be examining a system with a fixed temperature, volume, and number of particles in which case it is the Helmholtz energy as a functional of the number density profile $F[n(\vec r)]$ that is of interest. 
\color{black}

%*********************** DFT minimize F/N *********************
--------------------------------------------------------
\color{red}
Starting with Euler's equation
\begin{equation}U=TS-PV+\mu{N}\end{equation}
\begin{equation}\frac{U}{N}=\frac{TS}{N}-\frac{PV}{N}+\mu\end{equation}
\begin{equation}\label{d(U/N)long}d\left(\frac{U}{N}\right)=\frac{T}{N}dS+\frac{S}{N}dT+TSd\left(\frac{1}{N}\right)-\frac{V}{N}dP-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)+d\mu\end{equation}
and using the Gibbs-Duhem relation which follows from Euler's equation
\begin{equation}SdT-VdP+Nd\mu=0\end{equation}
which gives
\begin{equation}\frac{S}{N}dT-\frac{V}{N}dP+d\mu=0\end{equation}
Equation~\ref{d(U/N)long} becomes
\begin{equation}\label{d(U/N)}d\left(\frac{U}{N}\right)=\frac{T}{N}dS+TSd\left(\frac{1}{N}\right)-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
%\begin{equation}dU=TdS+SdT-PdV-VdP+\mu{dN}+Nd\mu\end{equation}
\color{black}
The Helmholtz free energy F is defined as 
\begin{equation}F=U-TS\end{equation}
\begin{equation}\frac{F}{N}=\frac{U}{N}-\frac{TS}{N}\end{equation}
\begin{equation}\label{usetoshowmin}d\left(\frac{F}{N}\right)=d\left(\frac{U}{N}\right)-d\left(\frac{TS}{N}\right)\end{equation}
\begin{equation}d\left(\frac{F}{N}\right)=d\left(\frac{U}{N}\right)-\frac{T}{N}dS-\frac{S}{N}dT-TS d\left(\frac{1}{N}\right)\end{equation}
After plugging in Equation~\ref{d(U/N)}, and setting $f_N=\frac{F}{N}$ this becomes
\begin{equation}df_N=-\frac{S}{N}dT-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
This result shows that the natural variables of the Helmholtz free energy per particle $f_N$ are temperature, volume and one over the number of particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed.
\begin{equation}f_N(T,V, \frac{1}{N})\end{equation}
Thus the Helmholtz free energy will be minimized in a system that has a temperature, volume, and number of particles that are fixed.

The minimization can be shown by considering that according to the second law of thermodynamics, a system that is not in equilibrium will expreience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$. Here "surroundings" means the entire universe minus the system. This, together with the conservation of energy (the first law of thermodynamics) which says that whatever energy is lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that any change in the free energy will be negative with each spontaneous process until a minimum is reached if the natural variables of the free energy are held fixed.  How will a system at a fixed temperature, volume, and number of particles respond to a small offset from equilibrium? Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
\begin{equation}d\left(\frac{F}{N}\right)_{System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\frac{S}{N}dT_{System}\end{equation}
and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that $dT=0$ gives
\begin{equation}d\left(\frac{F}{N}\right)_{System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}\end{equation}
where conservation of energy was also applied. Using Eq~\ref{d(U/N)} for the surroundings, but fixing the volume and the number of particles so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$ results in
\begin{equation}d\left(\frac{F}{N}\right)_{System}=-\frac{T}{N}dS_{System+Surroundings} \end{equation}which, for fixed N, is the same as
\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number of particles are fixed, the change in the free energy of the system will always be negative. (reference schroder)

More generally, the Helmholtz free energy is the value assumed by U-TS for a system that has a fixed volume, number of particles, and temperature that equals the temperature of the reservoir, or surroundings as it settles spontaneously to equilibrium. Under these conditions, the quantity U-TS settles to a minimum at equilibrium with respect to unconstrained variables, such as the number density profile $n(\vec r)$.
\begin{equation}F=\text{min}_{n(\vec r)}(U_{Not.equilibrium} - TS_{Not.equilibrium}) = U_{equilibrium}-TS_{equilibrium}\end{equation} An extremum is reached when 
\begin{equation}\frac{\partial{(U-TS)}}{\partial{n(\vec r)}}=0\end{equation}
Taking this derivative gives  \color{red} CHECK PARTIAL DERVIVATIVE OF FUNCTIONAL \color{black}
\begin{equation}\frac{\partial{(U-TS)}}{\partial{n(\vec r)}}=\frac{\partial{U}}{\partial{n(\vec r)}}-T\frac{\partial{S}}{\partial{U}}\bigg|_{V,N}\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
\begin{equation}=\frac{\partial{U}}{\partial{n(\vec r)}}-T\left(\frac{1}{T}\right)\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
\begin{equation}=\frac{\partial{U}}{\partial{n(\vec r)}}-\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
\begin{equation}=0\end{equation}
As shown earlier, the extremum that is reached at equilibrium is a minimum. (reference shethna) \color{red} To show that there is only one minimum, that the extremum is unique, I refer the reader to ...(reference fluids). \color{black} It remains to be shown that the Helmholtz free energy can be expressed as a functional of the number density profile.

------------------------------------------------------
%*************************END DFT minimize F/N ****************

@@

The Helmholtz free energy can be written as
\begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
where the paritian function Z is integrated over all of phase space for N identical particles each with a momentum $\vec{p}$ and a position $\vec{r}$:
\begin{equation}{Z=}\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dPdQ}\exp^\frac{-H(\vec{p}_1,\vec{p}_2,...\vec{p}_N,\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\end{equation}
\begin{displaymath}{dP=d\vec{p}_1d\vec{p}_2...d\vec{p}_N \mbox{~~~~where~~~~} d\vec{p}=dp_xdp_ydp_z}\end{displaymath}
\begin{displaymath}{dQ=d\vec{r}_1d\vec{r}_2...d\vec{r}_N \mbox{~~~~where~~~~} d\vec{r}=dxdydz\mbox{~~~~}}\end{displaymath}

Breaking the Hamiltonians H into its component energies:
\begin{equation}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
where the translational kinetic energy of the N atoms is given by: \begin{displaymath}{\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{displaymath}
the potential energy of the total interaction of each atom with the other N-1 atoms is given by:
\begin{displaymath}{V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} and \begin{displaymath}{V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} is an externally applied potential that varies spatially, such that its effects on each atom depend on the atom's position. Often $V_{ext}$ is set to zero, as it is in this paper.

Noting that the kinetic energy depends only on momentum while the total interatomic and external potential energies depend only on position, the integral in equation (8) can be separated into two parts: 
\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}\exp^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}\exp^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}  Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume defined by: \begin{equation}{V^N=}\int{...}\int{dQ}\end{equation} 
the equation can be written:
\begin{equation}{Z=\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}\exp^\frac{-(\frac{|\vec{p}_2|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)\left(\frac{1}{V^N}\int{...}\int{dQ}\exp^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}\right)
}\end{equation}  and the partitian function can be expressed as:
\begin{equation}\label{Zmultiply}{Z=Z_{ideal}Z_{excess}}\end{equation}
where:
\begin{equation}{Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}\exp^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}}\end{equation}
\begin{equation}{Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}\exp^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}}\end{equation} The ideal partitian function $Z_{ideal}$ is associated with an ideal gas which exhibits no interaction between the atoms - a condition approached by a gas at low density where the atoms rarely encounter one another. The excess partitian function $Z_{excess}$ accounts for the interaction of all the atoms according to their interatomic potential. The free energy can also be broken into two parts using Eq~\ref{Zmultiply}:
%\begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
\begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
\begin{equation}{F=-k_{B}T\ln(Z_{ideal})-k_{B}T\ln(Z_{excess})}\end{equation}
\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation}

The ideal Helmholtz free energy is
\begin{equation}{F_{ideal}= -k_BT\ln{\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}\exp^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)}}\end{equation} where 
%the integral can easily be evaluated:

>>>>

Another way to find values for $\Xi(T)$ is to use the second virial coefficent $B_{2}$ from the equation of state expressed in terms of the virial expansion. 
\begin{equation}\label{virialequation}
    \frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) 
\end{equation}
where 
\begin{equation}
	B_{2}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}(r)}-1\right)d\vec{r} 
\end{equation}
The second virial coefficient can be derived from Euler's equation for the internal energy of a system, 
\begin{equation}U=TS-PV+\mu{N}\end{equation}
using the relations from statistical thermophysics which views the thermodynamic quantities as ensemble averages
\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
which, for a canonical ensemble, gives
\begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
Expressing the Helmholtz free energy in terms of the pair potential $V(r)$ results in the equation of state expressed in terms of the virial expansion as given by Equation~\ref{virialequation}.

@@

\begin{figure}[h!]
    \centering
     %\includegraphics[height=7cm]{MaxwellDTC-Fig1.pdf}
     % FIXME\includegraphics[height=7cm]{plot-pressure_Fig1.pdf}
    \caption{Identifying densities at the solid-liquid transition using Maxwell's Double Tangent Construction}
    \label{fig:MaxwellDT}
  \end{figure}

@@

The pressure at the transistion point can be obtained from the point of intersection that occurs when the Gibbs free energy per atom is plotted against the pressure for both the liquid and the solid at fixed temperature, as shown in Figure \ref{fig:GibbsvsP}. These curves are constructed from the slope and cooresponding y-intercept data collected from lines tangent along the curves on the the Helmholtz free energy per atom verses inverse density plots for both the liquid and the solid. 
\begin{figure}[h!]
    \centering
     %\includegraphics[height=7cm]{MaxwellDTC-Fig2.pdf}
     %FIXME\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
    \label{fig:GibbsvsP}
  \end{figure}
%CURVE FITTING! - put in methods/analysis

@@

\begin{figure}[h!]
    \centering
    %\includegraphics[height=7cm]{MaxwellDTC-Fig3.pdf}
    % FIXME\includegraphics[height=7cm]{plot-pressure_Fig3.pdf}
    \caption{The liquid and solid densities at the point of transistion are found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
    \label{fig:Pvsinvn}
  \end{figure}

@@

A description of cDFT ordinarily begins with considering the Grand Free Energy defined as
\begin{equation}\Phi=F-\mu{N}\end{equation}
%\begin{equation}\int\frac{N}{V}\frac{V_{ext}}{N}dV\end{equation}
which can then be expressed in terms of a varrying number density profile $n(\vec r)$
%\begin{equation}\Phi=F-\int n(\vec{r})\phi{(\vec r)}d\vec{r} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu{N}\end{equation}
\begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
where $\phi=\frac{V_{ext}}{N}$, $V_{ext}(\vec r)$ is an externally applied potential energy, and
\begin{equation}F_{int}= F -\int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}  
%If $\mu=0$ and $V_{ext}=0$, then the Grand free energy $\Phi$ reduces to the Helmholz free energy $F$. %The second and third terms in Eq~\ref{GrandFE} are both functionals of $n(\vec{r})$, so it remains to show that $F$, and thus $\Phi$, is a functional of $n(\vec{r})$.

The Grand free energy is applicable to systems with a varying number of particles and fixed chemical potential, temperature and volume. But this paper will be examining a system with a fixed temperature, volume, and number of particles in which case it is the Helmholtz energy as a functional of the number density profile $F[n(\vec r)]$ that is of interest. 
%*********************** DFT minimize F/N *********************
%--------------------------------------------------------
%\color{red}
%Starting with Euler's equation
%%\begin{equation}U=TS-PV+\mu{N}\end{equation}
%\begin{equation}\frac{U}{N}=\frac{TS}{N}-\frac{PV}{N}+\mu\end{equation}
%\begin{equation}\label{d(U/N)long}d\left(\frac{U}{N}\right)=\frac{T}{N}dS+\frac{S}{N}%dT+TSd\left(\frac{1}{N}\right)-\frac{V}{N}dP-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)%+d\mu\end{equation}
%and using the Gibbs-Duhem relation which follows from Euler's equation
%\begin{equation}SdT-VdP+Nd\mu=0\end{equation}
%which gives
%\begin{equation}\frac{S}{N}dT-\frac{V}{N}dP+d\mu=0\end{equation}
%Equation~\ref{d(U/N)long} becomes
%\begin{equation}\label{d(U/N)}d\left(\frac{U}{N}\right)=\frac{T}{N}dS+TSd\left(\frac{1}{N}\right)-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
%\begin{equation}dU=TdS+SdT-PdV-VdP+\mu{dN}+Nd\mu\end{equation}
%\color{black}
The Helmholtz free energy F is defined as 
\begin{equation}F=U-TS\end{equation}
\begin{equation}\frac{F}{N}=\frac{U}{N}-\frac{TS}{N}\end{equation}
\begin{equation}\label{usetoshowmin}d\left(\frac{F}{N}\right)=d\left(\frac{U}{N}\right)-d\left(\frac{TS}{N}\right)\end{equation}
\begin{equation}d\left(\frac{F}{N}\right)=d\left(\frac{U}{N}\right)-\frac{T}{N}dS-\frac{S}{N}dT-TS d\left(\frac{1}{N}\right)\end{equation}
Setting $f_N=\frac{F}{N}$ this becomes
\begin{equation}df_N=-\frac{S}{N}dT-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
This result shows that the natural variables of the Helmholtz free energy per particle $f_N$ are temperature, volume and one over the number of particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed. It is the Helmholtz free energy will be minimized in a system that has a temperature, volume, and number of particles that are fixed.
\begin{equation}f_N(T,V, \frac{1}{N})\mbox{~~~~or~~~~}f_N(T,\frac{V}{N})\end{equation}

The minimization can be shown by considering that according to the second law of thermodynamics, a system that is not in equilibrium will expreience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$. Here "surroundings" means the entire universe minus the system. This, together with the conservation of energy (the first law of thermodynamics) which says that whatever energy is lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that any change in the free energy will be negative with each spontaneous process until a minimum is reached if the natural variables of the free energy are held fixed.  How will a system at a fixed temperature, volume, and number of particles respond to a small offset from equilibrium? Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
\begin{equation}d\left(\frac{F}{N}\right)_{System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\frac{S}{N}dT_{System}\end{equation}
and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that $dT=0$ gives
\begin{equation}d\left(\frac{F}{N}\right)_{System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}\end{equation}
where conservation of energy was also applied. Fixing the volume and the number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, results in
\begin{equation}d\left(\frac{F}{N}\right)_{System}=-\frac{T}{N}dS_{System+Surroundings} \end{equation}which, for fixed N, is the same as
\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number of particles are fixed, the change in the free energy of the system will always be negative. (reference schroder)

More generally, the Helmholtz free energy is the value assumed by U-TS for a system that has a fixed volume, number of particles, and temperature that equals the temperature of the reservoir, or surroundings as it settles spontaneously to equilibrium. Under these conditions, the quantity U-TS settles to a minimum at equilibrium with respect to unconstrained variables, such as the number density profile $n(\vec r)$.
\begin{equation}F=\text{min}_{U}(U_{Not.equilibrium} - TS_{Not.equilibrium}) = U_{equilibrium}-TS_{equilibrium}\end{equation}
An extremum is reached when
\begin{equation}\frac{\partial{F}}{\partial{U}}=0\end{equation}
\color{red} 
\begin{equation}F=\text{min}_{n(\vec{r})}(U_{Not.equilibrium} - TS_{Not.equilibrium}) = U_{equilibrium}-TS_{equilibrium}\end{equation} 
\color{red} 
\begin{equation}\frac{\delta{F}}{\delta{n(\vec r)}}\bigg|_{n(\vec{r})_{equil}}=0\end{equation}
Taking this derivative gives  ??????? CHECK PARTIAL DERVIVATIVE OF FUNCTIONAL 
%\begin{equation}\frac{\partial{(U-TS)}}{\partial{n(\vec r)}}=\frac{\partial{U}}{\partial{n(\vec r)}}-T\frac{\partial{S}}{\partial{U}}\bigg|_{V,N}\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
\begin{equation}\frac{\partial{(U-TS)}}{\partial{n(\vec r)}}=\frac{\partial{U}}{\partial{n(\vec r)}}-T\frac{\partial{S}}{\partial{U}}\bigg|_{V,N}\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
\begin{equation}=\frac{\partial{U}}{\partial{n(\vec r)}}-T\left(\frac{1}{T}\right)\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
\begin{equation}=\frac{\partial{U}}{\partial{n(\vec r)}}-\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
\begin{equation}=0\end{equation}
\color{black}
As shown earlier, the extremum that is reached at equilibrium is a minimum. (reference shethna) \color{red} To show that there is only one minimum, that the extremum is unique, I refer the reader to ...(reference fluids). \color{black} It remains to be shown that the Helmholtz free energy can be expressed as a functional of the number density profile.

%------------------------------------------------------
%*************************END DFT minimize F/N ****************

@@

The Helmholtz free energy can be written as
\begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
where the paritian function Z is integrated over all of phase space for N identical particles each with a momentum $\vec{p}$ and a position $\vec{r}$
\begin{equation}{Z=}\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dPdQ}~e^\frac{-H(\vec{p}_1,\vec{p}_2,...\vec{p}_N,\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\end{equation}
\begin{displaymath}{dP=d\vec{p}_1d\vec{p}_2...d\vec{p}_N \mbox{~~~~where~~~~} d\vec{p}=dp_xdp_ydp_z}\end{displaymath}
\begin{displaymath}{dQ=d\vec{r}_1d\vec{r}_2...d\vec{r}_N \mbox{~~~~where~~~~} d\vec{r}=dxdydz\mbox{~~~~}}\end{displaymath}

Breaking the Hamiltonians H into its component energies
\begin{equation}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
where the translational kinetic energy of the N atoms is given by \begin{displaymath}{\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{displaymath}
the potential energy of the total interaction of each atom with the other N-1 atoms is given by
\begin{displaymath}{V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} and \begin{displaymath}{V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} is an externally applied potential that varies spatially, such that its effects on each atom depend on the atom's position. Often $V_{ext}$ is set to zero, as it is in this paper.

Noting that the kinetic energy depends only on momentum while the total interatomic and external potential energies depend only on position, the integral in equation (8) can be separated into two parts 
\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}  Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume satisfying \begin{equation}{V^N=}\int{...}\int{dQ}\end{equation} 
%the equation can be written
%\begin{equation}{Z=\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}e^\frac{-(\frac{|\vec{p}_2|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)\left(\frac{1}{V^N}\int{...}\int{dQ}e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}\right)}\end{equation}  
and the partitian function can be expressed as
\begin{equation}\label{Zmultiply}{Z=Z_{ideal}Z_{excess}}\end{equation}
where
\begin{equation}{Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}}\end{equation}
\begin{equation}{Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}}\end{equation} The ideal partitian function $Z_{ideal}$ is associated with an ideal gas which exhibits no interaction between the atoms - a condition approached by a gas at low density where the atoms rarely encounter one another. The excess partitian function $Z_{excess}$ accounts for the interaction of all the atoms according to their interatomic potential. The free energy can also be broken into two parts using Equation~\ref{Zmultiply} giving
%\begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
\begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
\begin{equation}{F=-k_{B}T\ln(Z_{ideal})-k_{B}T\ln(Z_{excess})}\end{equation}
\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation}

The ideal Helmholtz free energy is
\begin{displaymath}{F_{ideal}=-k_BT\ln{\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)}}\end{displaymath}

\begin{equation}{\label{Fidealmultipleint}}{=-k_BT\ln\left(\frac{1}{h^{3}}\int{d\vec{p}_1}~e^\frac{-\frac{|\vec{p}_1|^2}{2m}}{k_BT}\frac{1}{h^{3}}\int{d\vec{p}_2}~e^\frac{-\frac{|\vec{p}_2|^2}{2m}}{k_BT}{...}\frac{1}{h^{3}}\int{d\vec{p}_N}~e^\frac{-\frac{|\vec{p}_N|^2}{2m}}{k_BT}\right)}\end{equation}
These integrals can conveniently be written in terms of the DeBroglie wavelength $\Lambda$
\begin{equation}{\frac{1}{h^{3}}\int{d\vec{p}}~e^\frac{-\frac{|\vec{p}|^2}{2m}}{k_BT}=\left(\sqrt{\frac{k_BTm}{2\pi\hbar^2}}\right)^3=\frac{1}{\Lambda^{3}}}\end{equation} 
where \begin{equation}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{equation} 
Rewriting Equation~\ref{Fidealmultipleint} in terms of $\Lambda$ gives
\begin{equation}{F_{ideal}= -k_BT[\ln(V^N)-\ln(N!) - \ln(\Lambda^{3N})]}\end{equation}Using Stirling's approximation \begin{displaymath}{\ln(N!)=N\ln{N}-N}\end{displaymath} this becomes
\begin{equation}{F_{ideal}= k_BT[N\ln(\frac{N\Lambda^{3}}{V}{)-N}]}\end{equation} 
Multiplying by $\frac{V}{V}$ gives an expression in terms of the number density $n=\frac{N}{V}$ gives \begin{equation}{F_{ideal}= k_BT\left[\frac{N}{V}\ln{\left(\frac{N\Lambda^3}{V}\right)}-\frac{N}{V}\right]}V\end{equation}
\begin{equation}{F_{ideal}= k_BT[n\ln(n\Lambda^3)-n]V}\label{eq:Fideal}\end{equation}   
For a spatially varying number density, the ideal free energy in Equation~\ref{eq:Fideal} becomes
\begin{equation}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]d\vec{r}}\end{equation}
